---
title: "R Plots Rutilus"
output: html_notebook
---


```{r setup, include=FALSE}

## Package Loading ## 
# for each package, check if it's installed and if not, install it

load_packages <- function(package_name)
  {
    if (!require(package_name, character.only = TRUE))
    {
      install.packages(package_name)
      library(package_name, character.only = TRUE)
    }
    else
    {
      library(package_name, character.only = TRUE)
    }
        
  }


Packages = c('dplyr', 'data.table', 'magrittr', 'ggplot2', 'ggpubr', 'plotly', 'plyr', 'tidyr', 'cowplot') 
invisible(lapply(Packages, load_packages))


# Setup knitr defaults with standard formatting changes
knitr::opts_chunk$set(
	echo = T,
	message = F,
	eval = T,
	fig.show = "markup",
	include = T,
	warning = F,
	results = "markup",
	tidy = T             
)

## Empty workspace ##
rm(list=ls())

new_theme <- theme_bw() + theme (
         panel.grid = element_blank(),
         plot.title = element_text(hjust = 0.5, vjust = 0, size = 12), #place title in the center and ajust the size
         plot.subtitle = element_text(hjust = 0.5, size = 10),#place subtitle in the center
         axis.title = element_text(size = 14), 
         axis.text = element_text(size = 14),
         axis.title.y = element_text(margin = margin(t = 0, r = 8, b = 0, l = 0)), #place title of the y axis not that close to the numbers
         axis.ticks = element_blank(),#remove ticks
         panel.border = element_blank(), #remove panel border 
         axis.line = element_line(size = 0.01, color ='#A9A9A9' ), #make axis thin and gray
         legend.position="right", 
         legend.text = element_text(size=8.5), #adjust legend text
         legend.title = element_text(size = 9), #adjust legend title
         legend.key.size = unit(0.8, "lines"), #adjust spacing between legend items - crucial if I want to make the whole legend smaller and not just the text
         plot.margin=unit(c(0.5,0.5,0.5,0.5),"cm"), #make margins so that all the legend is inside the plot area 
         aspect.ratio = 1/1.3)
theme_set(new_theme)
```

# m parameter variation
## m parameter coverage boxplot

First, I want to plot the coverage for different values of m. For this, I am first going to use bash commands (grep, awk, sed, cat) and then read the table into R and use ggplot. 

coverage_tables_m.sh script: 

```
##!/bin/bash

for m in 1 2 3 4 5 6 7 8 9 10
do
	cat stacks45.m.summary/stacks45.m$m/denovo_map.log | egrep -i '^Sample [0-9]|Stack coverage: mean=|Final coverage: mean='| sed "s/'/=/g" |  sed "s/;/=/g"| awk '{print $2}' FS='=' | xargs -n3 -d'\n' | sed "s/ /,/g" | sed -e "s/^/$m,/" > ./coverage_table_m$m.csv
done
cat coverage_table_m*.csv > coverage_table_m.csv
rm coverage_table_m?*.csv
```


```{r m parameter coverage create tables}
# running bash script from R studio
# the same can be done from command line
coverage_tables_m <- 'bash /mnt/c/Users/emeli/Documents/Rutilus_analysis/tests.denovo/coverage_tables_m.sh'
system(coverage_tables_m)
```

```{r m parameter coverage read table}
m_table <- data.table::fread("./coverage_table_m.csv")%>% 
  # make the variables factors 
  dplyr::mutate_at(
    vars(V1,V2),      
    list(factor)     
    )%>%
  # rename columns
  dplyr::rename(
    m = V1, 
    sample = V2, 
    initial_cov = V3, 
    final_cov = V4
    )
# reshape table into long 
m_table%<>%gather(., "cov_type", "coverage", 3:4)

```


```{r m parameter coverage boxplot}
colors = c("#6FBF55", "#9C89BD")
popmap = data.table::fread("./test_popmap.tsv")
n_samples = popmap %>% nrow()

m_cov_box<- ggplot(m_table, aes(y=coverage, x=m)) +
  geom_boxplot(
    alpha=0.2, 
    outlier.shape = NA, 
    aes(fill = cov_type), 
    position=position_dodge(0.6), 
    width = 0.4)+
  geom_point(
    position = position_jitterdodge(0.3), 
    alpha = 0.8, 
    size = 0.8, 
    aes(fill = cov_type, color = cov_type)) +
  theme (
    legend.position="none", 
    aspect.ratio = 1/2, 
    panel.grid.major.y = element_line(size = 0.1, color = '#d3d3d3'))+  
  labs(y = "coverage", 
       x = 'm',
       title = 'Coverage from denovo_map Rutilus, "Stack coverage" and "Final coverage"', 
       subtitle = paste("M=2, n=0,", n_samples, "samples from all populations")) +
  scale_fill_manual(values = colors)+
  scale_color_manual(values = colors)+
  scale_y_continuous(breaks = c(0, 5, 10, 15, 20, 25, 30, 35, 40, 45, 50, 55, 60, 65, 70), limits=c(0,75))

m_cov_box

ggsave(filename = './pics/fig1_boxplot_m_coverage.png', plot = m_cov_box,
    width = 20, height = 13, units = c("cm"),
    dpi = 300, limitsize = FALSE)
```


Now I want to plot snp per locus and total number of polymorphic loci. 

```
##!/bin/bash

for m in 1 2 3 4 5 6 7 8 9 10
do
	cat stacks45.m.summary/stacks45.m$m/populations.r80/populations.log.distribs| sed -n '/^BEGIN snps_per_loc_postfilters/,${p;/^END/q}' | sed '1d;2d;3d;$d'| sed -e "s/^/$m\t/"  > ./snps_per_locus_m$m.tsv
done
cat snps_per_locus_m*.tsv > snps_per_locus_m.tsv
rm snps_per_locus_m?*.tsv
```


```{r m parameter snp create tables}
# running bash script from R studio
# the same can be done from command line
snps_per_locus_m <- 'bash /mnt/c/Users/emeli/Documents/Rutilus_analysis/tests.denovo/snps_per_locus_m.sh'
system(snps_per_locus_m)
```


## m parameter snp per locus barplot
```{r m parameter snp per locus read table}
m_snp_table <- data.table::fread("./snps_per_locus_m.tsv")%>% 
  # rename columns
  dplyr::rename(
    m = V1, 
    n_snps = V2, 
    n_loci = V3
    )%>% 
  # make the variables factors 
  dplyr::mutate_at(
    vars(m, n_snps),      
    list(factor)     
    )
```

```{r m parameter snp per locus barplot}
library("viridis")          
colors = viridis(10)
m_snp_bar <- ggplot(m_snp_table , aes(y=n_loci/1000, x=n_snps, fill=m)) +
  geom_col(position=position_dodge(0.8))+  #position=position_dodge(0.8), width = 0.6
  labs(y = "n_loci (K)", 
       x = 'n_snps',
       title = 'snps per locus postfilters, Rutilus, from "populations.log.distribs"', 
       subtitle = paste("M=2, n=0,", n_samples, "samples from all populations"))+
  scale_fill_manual(values = colors )+
  scale_y_continuous(breaks = c(0, 2, 4, 6, 8, 10, 15, 20, 25, 30, 35, 40, 45, 50, 55)) +
  theme (panel.grid.major.y = element_line(size = 0.001, color = '#d3d3d3'), 
         axis.text = element_text(size = 9))
  
  m_snp_bar

ggsave(filename = './pics/fig2_barplot_m_snp_per_locus.png', plot = m_snp_bar ,
    width = 20, height = 13, units = c("cm"),
    dpi = 300, limitsize = FALSE)
```


Now I want to plot all assembled loci: the sum of all n_loci per one value of m. 

## m parameter n_loci_assembled
```{r m parameter n_loci table}
# reading the same table again so that this chunk could be executed separately 
n_loci_assembled_table <- data.table::fread("./snps_per_locus_m.tsv")%>% 
  # rename columns
  dplyr::rename(
    m = V1, 
    n_snps = V2, 
    n_loci = V3
    )%>% 
  # make the variables factors 
  dplyr::mutate_at(
    vars(m, n_snps),      
    list(factor)     
    )%>%
  # group by the factor m 
  dplyr::group_by(m)%>%
  # count the sum of the loci column, including even those loci where no snps were found 
  dplyr::summarise(
      n_loci_assembled = sum (n_loci)
    )

```

To count the number of polymorphic sites, we should do the following, but without m = 0 

```{r n loci polymorphic table }

n_loci_polymorphic_table <- data.table::fread("./snps_per_locus_m.tsv")%>% 
  # rename columns
  dplyr::rename(
    m = V1, 
    n_snps = V2, 
    n_loci = V3
    )%>% 
  # make the variables factors 
  dplyr::mutate_at(
    vars(m, n_snps),      
    list(factor)     
    )%>%
  #filter out the cases where loci have no sps 
  dplyr::filter(n_snps!=0)%>%
  # group by the factor m 
  dplyr::group_by(m)%>%
  # count the sum of the number of loci column
  dplyr::summarise(
      n_loci_polymorphic = sum (n_loci)
    )

# join 2 tables: 

m_loci_table <- left_join(n_loci_assembled_table  , n_loci_polymorphic_table, by = c('m'))


```
Plotting the values of assembled and polymorphic loci. 

```{r m parameter n_loci_assembled and polymorphic dot plot }

m_loci_ass_dot <- ggplot(m_loci_table, aes(y=n_loci_assembled/1000, x=m)) +
  geom_point(size = 3) +
  geom_point(size = 3, color = '#6B8E23', aes(y=n_loci_polymorphic/1000, x=m))+
  theme (legend.position="none", panel.grid.major.y = element_line(size = 0.1, color = '#d3d3d3'))+  #remove legend
  labs(y = "n_loci_assembled (K)", 
       x = 'm',
       title = 'Number of assembled loci from populations.log.distribs', 
       subtitle = "M=2, n=0, r80, green - polymorphic loci") +
  scale_y_continuous(labels = scales::comma, breaks = c(0, 10, 20, 30, 40, 50, 60, 70), limits=c(0,70))
m_loci_ass_dot

ggsave(filename = './pics/fig3_dotplot_m_assembled_polymorphic_loci.png', plot = m_loci_ass_dot,
    width = 20, height = 12, units = c("cm"),
    dpi = 300, limitsize = FALSE)

ggplotly(m_loci_ass_dot)

```

## m parameter number of SNPs 

```{r n loci polymorphic table }

n_snps_table <- data.table::fread("./snps_per_locus_m.tsv")%>% 
  # rename columns
  dplyr::rename(
    m = V1, 
    n_snps = V2, 
    n_loci = V3
    )%>% 
  # make the variables factors 
  dplyr::mutate_at(
    vars(m),      
    list(factor)     
    )%>%
  #filter out the cases where loci have no sps 
  dplyr::filter(n_snps!=0)%>%
  dplyr::mutate(
    number_of_snp = n_snps*n_loci
  )%>%
  # group by the factor m 
  dplyr::group_by(m)%>%
  # count the sum of the number of snp column
  dplyr::summarise(
      n_snp = sum (number_of_snp)
    )



```

```{r m parameter n_snps dot plot }

m_snps_dot <- ggplot(n_snps_table, aes(y=n_snp/1000, x=m)) +
  geom_point(size = 3) +
  theme (legend.position="none", panel.grid.major.y = element_line(size = 0.1, color = '#d3d3d3'))+  #remove legend
  labs(y = "n_snps (K)", 
       x = 'm',
       title = 'Number of snps from populations.log.distribs', 
       subtitle = "M=2, n=0, r80") +
  scale_y_continuous(labels = scales::comma, breaks = c(0, 10, 20, 30, 40, 50, 60, 70), limits=c(0,70))
m_snps_dot

ggsave(filename = './pics/fig4_dotplot_m_snps.png', plot = m_snps_dot,
    width = 20, height = 12, units = c("cm"),
    dpi = 300, limitsize = FALSE)

```